---
output: 
  github_document:
    toc: false
    fig_width: 10.08
    fig_height: 6
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
```

# `ggggeffects`

The goal of `ggggeffects` is to allow for a `ggplot2` syntax for the actual plotting of the data-grid provided by `ggeffects`.

I built this because I love `ggeffects`, but I wanted more control over the plotting.

```{r load_ggge}
library(ggggeffects)
```

## Installation

```r
# install.packages("remotes")
remotes::install_github("mattansb/ggggeffects")
```

## Examples

Fit a model from the `mtcars` data:

```{r mtcars_model}
fit <- lm(mpg ~ factor(cyl) + am + disp + hp, mtcars)
```

### Full `ggplot2` Customizability

```{r compare}
gge <- ggpredict(fit, terms = c("cyl", "am"))

# Compare
plot(gge)

# To:
autoplot(gge) +
  geom_CI_bar() +
  geom_expected_point()
```

The default plot is better looking than the ugly `ggplot2` default, however `ggggeffects` is fully customisable, as would be expected (also not that the `am` is retained as a numeric variable):

```{r custom1}
autoplot(gge, aes(color = factor(am), fill = after_scale(color)),
         cols = vars(am)) +
  geom_CI_bar(width = .3, color = "black", position = position_dodge(0.3)) +
  geom_expected_point(size = 2, position = position_dodge(0.3)) +
  theme_bw() +
  labs(color = "am")
```

Here is another example, with a continuous variable on x:

```{r custom2}
gge <- ggpredict(fit, c("disp", "hp", "am"))

(p <- autoplot(gge) +
  geom_CI_ribbon(color = NA, alpha = 0.2) +
  geom_expected_line() +
  scale_color_distiller(type = "div", palette = 7, 
                        aesthetics = c("color", "fill")) +
  theme_bw())
```

### Adding Raw/Residualized Data

We can use `stat_raw_data` to add the raw data to the plot! It's so simple!

```{r with_data}
p + stat_raw_data()
```

We can also create partial residual plots with `stat_residualized_data()`:

```{r with_residuals}
p + stat_residualized_data()
```

We can even plot them side by side:
```{r with_both}
p +
  stat_raw_data(aes(shape = "raw"), size = 2, alpha = 0.6) +
  stat_residualized_data(aes(shape = "residuals"), size = 2,
                         position = position_nudge(x = 10)) +
  labs(shape = "Data type")
```

### Raster Plots

We can also plot bi-variate raster plots:

```{r raster_2Dplot}
gge <- ggpredict(fit, c("disp [n=10]", "hp [n=10]", "am"))

autoplot(gge, aes(x = disp, y = hp,
                  color = predicted, fill = predicted)) +
  geom_raster() +
  # geom_contour(aes(z = predicted)) +
  stat_raw_data(shape = 21, color = "white", size = 2) +
  scale_color_viridis_c(aesthetics = c("color", "fill")) +
  theme_minimal()
```


### Collapsing Across Random Variables

Here is an example from MLM Stroop data:

```{r stroop}
library(dplyr)

data("stroop", package = "afex")

set.seed(42)

some_stroop_data <- stroop %>%
  filter(study == 1 & acc == 1, trialnum < 60) %>%
  sample_n(1000) %>%
  mutate(
    pno = factor(pno),
    level1_cov = rnorm(n())
  ) %>%
  group_by(pno) %>%
  mutate(level2_cov = rnorm(1)) %>%
  ungroup()

m <- lme4::lmer(rt ~ condition + congruency + level1_cov + level2_cov +
                  (congruency | pno) + ( 1 | trialnum),
                data = some_stroop_data)


gge <- ggemmeans(m, c("congruency", "condition"))


p <- autoplot(gge) +
  geom_CI_bar(width = .3, color = "black", position = position_dodge(0.7)) +
  geom_expected_point(shape = 21, color = "black", 
                      position = position_dodge(0.7), size = 3)
```

To collapse across a grouping variable, we set `collapse.group` to `TRUE`:

```{r collapsTRUE}
p +
  stat_raw_data(collapse.group = TRUE, 
                position = position_jitterdodge(0.1,,0.7), alpha = 0.3)
```

Or specify by name:

```{r collapse_trialnum}
p +
  stat_raw_data(collapse.group = "trialnum", 
                position = position_jitterdodge(0.1,,0.7), alpha = 0.3)
```

We can combine and mix as we please:

```{r collapse_both}
p +
  stat_raw_data(collapse.group = "pno", 
                aes(shape = "pno", x = as.numeric(congruency) + .05),
                position = position_jitterdodge(0.1,,0.7), alpha = 0.3) +
  stat_raw_data(collapse.group = "trialnum", 
                aes(shape = "trialnum", x = as.numeric(congruency) - .05),
                position = position_jitterdodge(0.1,,0.7), alpha = 0.3) +
  scale_shape_manual("Random var", values = c(16, 8))
```


We can do the same for residualized data, or with a continuous variable on x (though this only really works nicely if the covariate is on level 2+):

```{r}
gge <- ggemmeans(m, c("level2_cov [all]", "congruency", "condition"))

autoplot(gge) +
  geom_CI_ribbon(color = NA, alpha = 0.1) +
  geom_expected_line() +
  stat_residualized_data(collapse.group = "pno", alpha = 0.4)
```

